---
title: "TidyTuesdayGreatAmericanBeerFestival"
author: "Jeff Grayum"
date: "12/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries:
```{r}
library(tidyverse)
library(tidytuesdayR)
library(ggthemes)
library(tvthemes)
library(scales)
library(glue)
library(extrafont)
#install.packages("tidylo")
library(tidylo)
library(tidytext)
library(broom)
#install.packages("maps")
library(maps)
```

Loading data:
```{r}
tuesdata <- tidytuesdayR::tt_load('2020-10-20')

beer_awards <- tuesdata$beer_awards %>%
  mutate(state = str_to_upper(state),
         medal = fct_relevel(medal, c("Bronze", "Silver"))) 
  

gold_medals <- beer_awards %>%
  filter(medal == "Gold") %>%
  view()
```

Initial analysis and some visualizations:
```{r}
beer_awards %>%
  view()

beer_awards %>%
  distinct(year)

gold_medals %>%
  count(brewery, sort = TRUE) %>%
  head(20) %>%
  mutate(brewery = fct_reorder(brewery, n)) %>%
  ggplot(aes(n, brewery)) + 
  geom_col() +
  labs(title = "Which breweries have won the most gold medals at the Great American Beer Festival",
       subtitle = "Gold medals awarded from 1987 - 1920",
       x = "Total gold medals",
       y = "",
       caption = "Source: Great American Beer Festival") +
  theme_simpsons() 

beer_awards %>%
  count(brewery, sort = TRUE) %>%
  head(20) %>%
  mutate(brewery = fct_reorder(brewery, n)) %>%
  ggplot(aes(n, brewery)) + 
  geom_col() +
  labs(title = "The twenty most awarded breweries in the North American Beer Festival",
       subtitle = "Total medals awarded from 1987 - 1920",
       x = "Total medals",
       y = "",
       caption = "Source: Great American Beer Festival") +
  theme_simpsons() 

beer_awards %>%
  count(city, sort = TRUE) %>%
  head(20) %>%
  mutate(city = fct_reorder(city, n)) %>%
  ggplot(aes(n, city)) +
  geom_col() +
  theme_simpsons() +
  labs(title = "Which cities brew the best beer?",
       x = "Number of times a brewery has medaled at Great American Beer Festival",
       y = "",
       caption = "Source: Great American Beer Festival 1987 - 2020")

beer_awards %>%
  count(state, sort = TRUE) %>%
  head(20) %>%
  mutate(state = fct_reorder(state, n)) %>%
  ggplot(aes(n, state)) +
  geom_col() +
  theme_simpsons() +
  labs(title = "Which States brew the best beer?",
       x = "Number of times a brewery has medaled at Great American Beer Festival",
       y = "",
       caption = "Source: Great American Beer Festival 1987 - 2020")


beer_awards %>%
  count(beer_name, brewery, sort = TRUE) %>%
  head(20) %>%
  mutate(beer_name = fct_reorder(beer_name, n)) %>%
  ggplot(aes(n, beer_name)) +
  geom_col() +
  labs(title = "What are the most highly-awarded beers at the Great American Beer Festival?",
       x = "Number of medals won",
       y = "",
       caption = "Source: Great American Beer Festival 1987 - 2020") +
  theme_simpsons()

gold_medals %>%
  count(beer_name, brewery, sort = TRUE) %>%
  head(5) %>%
  mutate(beer_name = fct_reorder(beer_name, n)) %>%
  ggplot(aes(n, beer_name)) +
  geom_col() +
  labs(title = "Which beers have won the most gold medals at the Great American Beer Festival?",
       x = "Number of medals won",
       y = "",
       caption = "Source: Great American Beer Festival 1987 - 2020") +
  theme_simpsons()

beer_awards %>%
  group_by(year) %>%
  count(category, sort = TRUE) %>%
  View()

beer_awards %>%
  filter(category == "IPA") %>%
  count(beer_name, sort = TRUE) %>%
  head(5) %>%
  view()
  mutate(beer_name = fct_reorder(beer_name, n)) %>%
  ggplot(aes(n, beer_name)) +
  geom_col()
  
  beer_awards %>%
    filter(category == "American-Style Pale Ale") %>%
    count(brewery, sort = TRUE) %>%
    head(10) %>%
    mutate(brewery = fct_reorder(brewery, n)) %>%
    ggplot(aes(n, brewery)) +
    geom_col() +
    labs(title = "What brewery makes the best IPA?",
         x = "Number of medals",
         y = "", 
         caption = "Source: Great American Beer Festival 1987 - 2020") +
    theme_simpsons()
```

Okay, now for David's take.
```{r}
beer_awards %>%
  count(year) %>%
  ggplot(aes(year, n)) + 
  geom_col() +
  labs(title = "How many awards were handed out each year during the Great American Beer Festival?",
       x = "Year",
       y = "Number of medals awarded",
       caption = "Sourch: Great American Beer Festival 1987 - 2020") +
  theme_simpsons()

beer_awards %>%
  count(brewery, city, state, sort = TRUE) 

beer_awards %>%
  count(category, sort = TRUE)

beer_awards %>%
  filter(category == "Bock") %>%
  view()
```

What are the most awarded beers?
```{r}
beer_awards %>%
  count(beer_name, brewery, sort = TRUE)

beer_counts <- beer_awards %>%
  count(beer_name, brewery, medal, city, state) 

beer_counts %>%
  pivot_wider(names_from = medal, values_from = n,
              values_fill = list(n = 0)) %>%
  mutate(total = Bronze + Silver + Gold) %>%
  arrange(desc(total))
```

HERE'S A REALLY COOL TRICK ON HOW TO USE FILTER() AND FCT_LUMP() TO SHOW ONLY THE TWELVE MOST AWARDED BEERS!!
```{r}
beer_counts %>%
  mutate(beer_name = glue("{ beer_name } ({ brewery })")) %>%
  filter(fct_lump(beer_name, 20, w = n, ties.method = "first") != "Other")  %>%
  mutate(beer_name = fct_reorder(beer_name, n, sum)) %>%
  ggplot(aes(n, beer_name, fill = medal)) +
  geom_col() +
  labs(title = "The most awarded beers at the Great American Beer Festival",
       x = "Number of Awards",
       y = "",
       fill = "", 
       caption = "Source: Great American Beer Festival, 1987 - 2020") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman")) +
  scale_fill_manual(values = myColors)

myColors <- c("#a37d0b", "#9bc4cc", "#ebda21")
```

What breweries win the most awards?
```{r}
beer_awards %>%
  count(brewery, sort = TRUE)

by_brewery <- beer_awards %>%
  group_by(brewery) %>%
  summarize(n_medals = n(),
            n_beers = n_distinct(beer_name),
            weighted_medals = sum(as.integer(medal)), #medals are already factored
            n_gold = sum(medal == "Gold")) %>% 
  arrange(desc(n_medals)) %>%
  view()

beer_awards %>%
  count(brewery, medal) %>%
  filter(fct_lump(brewery, 16, w = n) != "Other") %>%
  mutate(brewery = fct_reorder(brewery, n, sum)) %>%
  ggplot(aes(n, brewery, fill = medal)) +
  geom_col() +
  theme_fivethirtyeight() +
  labs(title = "The most awarded breweries at the Great American Beer Festival",
       x = "Number of awards",
       y = "",
       caption = "Source: Great American Beer Festival, 1987 - 2020",
       fill = "") +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman")) +
  scale_fill_manual(values = myColors)
```

Now, let's look at the states!
```{r}
beer_awards %>%
  count(state, sort = TRUE)

# SET OF INTERNAL DATA CALLED STATE.ABB!!
state.abb

#We can compare it to our count of states using setdiff() to see if any are missing.
setdiff(state.abb, beer_awards$state)

#We see that West Virginia has never won an award.  Probably because this isn't a moonshine data set.

beer_awards %>%
  count(state, medal) %>%
  filter(fct_lump(state, 16, w = n) != "Other") %>%
  mutate(state = fct_reorder(state, n, sum)) %>%
  ggplot(aes(n, state, fill = medal)) +
  geom_col() +
  theme_fivethirtyeight() +
  labs(title = "The states recieving the most awards at the Great American Beer Festival",
       x = "Number of awards",
       y = "",
       caption = "Source: Great American Beer Festival, 1987 - 2020",
       fill = "") +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman")) +
  scale_fill_manual(values = myColors)

#Let's look at per capita award winning beers


```

TO DO
-Choropleth (colored map) 
-Per-capita award winning beers

```{r}
breweries <- read_csv("~/Downloads/archive/breweries.csv") %>%
  rename(brewery_id = X1,
         brewery = name) %>%
  view()  


beers <- read_csv("~/Downloads/beers.csv") %>%
  rename(beer_name = beer) %>%
  inner_join(breweries, by = "brewery_id") %>%
  view()

beers %>%
  view()

beers %>%
  mutate(style = fct_lump_n(style, 10)) %>% 
  filter(style != "NA") %>%
  ggplot(aes(abv, fill = style)) + 
  geom_histogram() +
  labs(title = "ABV distribution",
       x = "ABV",
       y = "",
       caption = "Source: Kraggle.com") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman")) +
  scale_x_continuous(labels = percent)

beers %>%
  filter(!is.na(style)) %>%
  count(style, sort = TRUE) %>%
  head(20) %>%
  mutate(style = fct_reorder(style, n)) %>%
  ggplot(aes(n, style)) +
  geom_col() +
  theme_fivethirtyeight() +
  labs(title = "The most popular styles of craft beer",
       x = "",
       y = "",
       caption = "Source: Kraggle.com") +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

breweries <- read_csv("~/Downloads/archive/breweries.csv") %>%
  rename(brewery_id = X1,
         brewery = name) %>%
  view()

breweries %>%
  view()

breweries %>%
  count(city, sort = TRUE) %>%
  head(20) %>%
  mutate(city = fct_reorder(city, n)) %>%
  ggplot(aes(n, city)) +
  geom_col() +
  theme_fivethirtyeight() +
  labs(title = "Which cities have the most microbreweries?",
       x = "",
       y = "") 

breweries %>%
  count(state, sort = TRUE) %>%
  head(20) %>%
  mutate(state = fct_reorder(state, n)) %>%
  ggplot(aes(n, state)) +
  geom_col() +
  theme_fivethirtyeight() +
  labs(title = "Which states have the most microbreweries?",
       x = "",
       y = "") 
  
beers %>%
  mutate(brewery = glue("{ brewery } ({ state })")) %>%
  count(brewery, sort = TRUE) %>%
  head(20) %>%
  mutate(brewery = fct_reorder(brewery, n)) %>%
  ggplot(aes(n, brewery)) +
  geom_col() +
  labs(title = "What craft breweries produce the greatest variety of beer?",
       x = "",
       y = "",
       caption = "Source: CraftCans.com") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(), 
        text = element_text(family = "Times New Roman"))

beers %>%
  filter(state == "AZ") %>%
  mutate(brewery = glue("{ brewery } ({ city })")) %>%
  count(brewery, sort = TRUE) %>%
  mutate(brewery = fct_reorder(brewery, n)) %>%
  ggplot(aes(n, brewery)) +
  geom_col() +
  labs(title = "The biggest breweries in my home state of Ari - fucking - zona",
       x = "",
       y = "",
       caption = "Source: CraftCans.com",
       subtitle = "Note: Mother Road is the best brewery in AZ and is not on this list") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"))

```


Let's look at a combination of category and state.
```{r}
#First filter for common states

beer_awards %>%
  filter(fct_lump(state, 12) != "Other",
         fct_lump(category, 10) != "Other") %>%
  count(state, category) %>%
  complete(state, category, fill = list(n = 0)) %>%
  bind_log_odds(state, category, n) %>%
  mutate(state = fct_reorder(state, -n, sum)) %>%
  mutate(category = reorder_within(category, log_odds_weighted, state)) %>%
  ggplot(aes(log_odds_weighted, category, fill = log_odds_weighted > 0)) +
  geom_col() +
  scale_y_reordered() +
  facet_wrap(~ state, scales = "free_y") +
  labs(title = "Representation of categories of beer in each state",
       subtitle = "Representation judged by awards recieved at the Great American Beer Festival",
       x = "Weighted log-odds",
       y = "",
       caption = "Source: CraftCans.com") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"),
        legend.position = "none")
```


Let's look at states over time!
```{r}
beer_awards %>%
  count(year, state, sort = TRUE)

beer_awards %>%
  add_count(year, name = "year_total")

by_year_state <- beer_awards %>%
  add_count(year, name = "year_total") %>%
  mutate(state = fct_lump(state, 9)) %>%
  count(year, state, year_total, sort = TRUE) %>%
  mutate(pct_year = n / year_total)

by_year_state %>%
  filter(state != "Other") %>%
  ggplot(aes(year, pct_year, color = state)) +
  geom_line() + 
  expand_limits(y = 0) +
  facet_wrap(~ state) +
  labs(title = "Percentage of awards won each year at the Great American Beer Festival",
       x = "",
       y = "",
       caption = "Source: CraftCans.com") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman"),
        legend.position = "none") +
  scale_y_continuous(labels = percent)

#There will be a lot of "noise" in early years, since we had fewer categories and less beers.  We can compensate for this poor data with logistic regression!!




by_year_state %>%
  filter(state == "CA") %>%
  glm(cbind(n, year_total - n) ~ year, 
      data = ., 
      family = "binomial") %>%
  summary()

by_year_state %>%
  filter(state != "Other") %>%
  mutate(state = state.name[match(state, state.abb)]) %>%
  group_by(state) %>%
  summarize(model = list(glm(cbind(n, year_total - n) ~year, family = "binomial")))%>%
  mutate(tidied = map(model, tidy, conf.int = TRUE)) %>%
  unnest(tidied) %>%
  filter(term == "year") %>%
  mutate(p.value = format.pval(p.value),
         state = fct_reorder(state, estimate)) %>%
  ggplot(aes(estimate, state)) +
  geom_point() +
  geom_vline(xintercept = 0, lty = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.1) +
  labs(x = "Estimated slope",
       title = "Which states become more ore less frequent medal winners?", 
       y = "",
       caption = "Source: CraftCans.com") +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        text = element_text(family = "Times New Roman")) 


```

A quick chloropleth (heat map) !!
```{r}
beer_awards %>%
  mutate(state = str_to_lower(state.name[match(state, state.abb)])) %>%
  count(state) %>%
  right_join(map_data("state"), by = c(state = "region")) %>%
  ggplot(aes(long, lat, fill = n, group = group)) +
  geom_polygon() +
  theme_map() +
  labs(fill = "Number of award winning beers",
       title = "A map of which states have been most successful at the Great American Beer Festival",
       caption = "Source: CraftCans.com") 
  
  